Efficient static analysis of Marlowe contracts

### 2020-05-20 papers/static_analysis2020/paper.tex
Add measurements section explanation

>>>>

In our experiments, the heavy weight optimisations considerably reduced
the requirements of both processing time and memory for Marlowe's static
analysis. In Tables~\ref{table1},~\ref{table2},~\ref{table3},~and~\ref{table4},
we present the results of measuring the performance of static analysis on
four example Marlowe contracts,
before\footnote{\url{https://github.com/input-output-hk/marlowe/blob/master/src/Language/Marlowe/Analysis/FSSemantics.hs} [last visited 19-05-2020]}
and after\footnote{\url{https://github.com/input-output-hk/marlowe/blob/master/src/Language/Marlowe/Analysis/FSSemanticsFastVerbose.hs} [last visited 19-05-2020]}
the heavyweight optimisations.

Unfortunately, at the time of writing, we do not have a completely unoptimised
version of the static analysis that we can use to compare, because the semantics
of Marlowe have changed since we implemented such a version.

Using the \texttt{perf} tool \cite{perf}, we have measured the execution time
and the overhead of both the generation of the constraints and their solution
by Z3 \cite{de2008z3}. And we have measured the peak RAM usage of the whole
process by using GNU's \texttt{time} tool \cite{keppel1998gnu}.

In all the contracts, the implementation that has the heavyweight optimisations
performs much better and scales further. In the case of the \textit{auction}
(Table~\ref{table1}) and \textit{crowdfunding} (Table~\ref{table2}) contracts,
whose size grows exponentially, both approaches quickly overwhelm the resources
available in the execution environment. On the other, in the case of \textit{rent}
(Table~\ref{table3}) and \textit{coupon bond} (Table~\ref{table4}), which grow
linearly, by using the lightweight version the problem becomes intractable much
faster than with the heavyweight version.

Another conclusion that can be derived from the experiments is that, in the
version with lightweight optimisations, most of the processing time seems to be
spent during the generation of the constraints, and solving is done relatively
quickly by Z3; while the opposite happens for the version with heavyweight
optimisations.

These results suggest that SBV library is able to generate constraints in a way
that they are handled efficiently by Z3, but the process itself can be costly
if we are not careful. Nevertheless, the overall execution time required by Z3
is also lower in the case of the version with heavyweight optimisations and it
grows more slowly, which suggests the optimisations described in this paper
affect both stages of the process.
